---
title: "Zombie HW"
author: "Laura Brubaker-Wittman"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework Set-up:

*Load in the dataset “zombies.csv” from my GitHub repo at
<https://github.com/fuzzyatelin/fuzzyatelin.github.io/tree/master/AN597_Fall19/>. 
This data includes the first and last name and gender of the entire population of 1000 people who have survived the
zombie apocalypse and are now ekeing out an existence somewhere on the East Coast, along with several other
variables (height, weight, age, number of years of education, number of zombies they have killed, and college
major).* 

In order to do this, I will run the code below, after loading the {curl} library:
```{r}
library(curl)
f <- f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall19/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

This shows me the first 6 rows of the data and my 'Global Environment' in R Studio also shows that this data is now
loaded for this work session.
